---
title: "Macro SSDI"
author: "Chauncey Gadek" 
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    number_sections: true
    toc_depth: 5
    code_folding: show
    bibliography: true
    #df_print: paged
    #df_print: kable
    #toc_float: true
      #collapsed: false
      #smooth_scroll: TRUE
    theme: cosmo #spacelab #yeti #united #cosmo
    highlight: tango
  pdf_document:
    df_print: kable
fontsize: 12pt
geometry: margin=0.25in
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

```{=html}
<style>
/* HTML FORMATTING */
h1, .h1, h2, .h2, h3, .h3, h4, .h4, h5, .h5 {
  margin-top: 25px; /* space before each header */
  font-weight: bold; /* bold headers */
}
</style>
```
```{R, echo=FALSE}
# I set some GLOBAL R chunk options here.
#   (to hide this message add "echo=FALSE" to the code chunk options)
rm(list =ls (all = TRUE)) #This removes objects from global environ
knitr::opts_chunk$set(echo=F, comment = NA, message = FALSE, warning = FALSE, width = 100)
knitr::opts_chunk$set(fig.align = "center", fig.height = 4, fig.width = 6)
#setwd("C:/Users/Trinity's/Dropbox/Crane_trachea")
#setwd("~/Dropbox/Crane_trachea") #need to change to local directory
```

# Load Packages

```{R, echo=F}
pacman::p_load(
reshape,
reshape2,
plyr,
dplyr,
tidyr,
car,
picante,
rcompanion,
GGally,
Hmisc,
gridExtra,
stats,
gplots,
ggplot2,
ggExtra,
cowplot,
colorspace,
stats4, # Forces knitr to work when it's being wonky
PMCMR, #Allows Kruskal-Wallis post-hocs
effects,
gridExtra,
lattice,
survival,
fmsb,
faraway,
ape,
data.table,
#wBoot,
ggridges,
boot,
faux,
effsize,
plotrix,
colorspace,
ggpubr,
patchwork,
ggdist,
factoextra,

# Mapping 
raster,
sp,
rgdal,
RStoolbox,
prettymapr,
viridis,
rasterVis,


# Modeling packages 
nlme,
lme4,
AICcmodavg,
MuMIn,
glmulti,
reghelper,
lsmeans,
rsq, # get r-squared values from GLM
r2glmm, # for R^2 values from lmer(, and glmer(,
multcompView, # related to multiple comparisons?
jtools, # interaction plots 
interactions, # interaction plots 
broom,
stargazer, # model output tables
ggeffects, # for estimating model predictions from mixed effects models
MCMCglmm,
bayesplot,
rstan,
Rcpp, # required for brms
brms,
magrittr,
tidybayes,
modelr,
hexbin,
ggExtra,
rgl,
readr,
tidyverse,
# Phylo packages 
phytools,
ape
)

# To run each time you load rstan
options(mc.cores = parallel::detectCores()) # for core setup 
rstan_options(auto_write = TRUE) # auto save  bare version of compiled Stan program to HD



#Load in functions
source("~/Desktop/R_color_palettes/Gadek_custom_colors.R")
source("~/Desktop/ggplot_themes/ggplot_themes.R")

#set theme
theme_set(theme_arial_clean())

#setup folder paths for less coding
figures <- paste(getwd(), "/figures/", sep="")
tables <- paste(getwd(), "/Tables/", sep="")
models <- paste(getwd(), "/models/", sep="")
results <- paste(getwd(), "/models/results/", sep="")
```

Sexual size dimorphism is a long-studied phenomenon in animals. In this exploration, I examine a few historical methods of effectively comparing and scaling SSD between species @lovich1992review, and I extend studies of sexual dimorphism across elevation gradients, mainly in insects,(@laiolo2013local, @bowden2013fecundity), to Andean birds using phylogenetic mixed models.

The debate about the proper metric for comparing sexual size dimorphism among taxa has yielded one heuristic: simpler is better. @lovich1992review reviewed various ratio metrics of sexual size dimorphism looking at scalability and intuitive utility and recommend: 

$$
SDI = \frac{size of larger sex}{size of smaller sex}
$$




I'm mainly using the dataset used in @linck2021blood downloaded from Ethan Linck's [Github repo](https://github.com/elinck/andean_range_limits). I am supplementing this with elevational range data collated from @del1992handbook and avian trait data downloaded from the [Open Traits Network](https://opentraits.org/datasets/avonet), (@tobias2022avonet). These have been processed in the MACRO_morpho_data_wrangling.R file. No need to rerun unless adding/adjusting data/data structure. 

Read in the processed dataframe here.

```{r}

df<-read_csv("data/All_birds_elev_macro_morpho.csv")

df<- df%>%
  mutate(sex_diff = mean_mass_male-mean_mass_female,
    sex.diff.corr = log10(sex_diff - min(sex_diff, na.rm=T)+1),
    mdpt.elev = (max_elev+min_elev)/2,
    amplitude = max_elev -min_elev,
    larger_sex = if_else(mean_mass_male > mean_mass_female, "male larger", "female larger"))

#tree <- read.tree(paste0("data/tree_2801spp.tre")) #need to trim to model. will take a million years to run with 9000+ tips anyway
```

# Explore data

```{r, `birdsexdiff`, warning=F}
df%>%
   filter(sex_diff> -2000,
          family %in% c("Tyrannidae", "Furnariidae", "Emberizidae", "Fringillidae", "Psittacidae", "Trochilidae"))%>%
ggplot(., aes(log10(mdpt.elev), sex.diff.corr))+
  #geom_point(aes(size=Range.Size), shape=21, alpha=0.4, fill="grey")+
  geom_smooth(method="lm", color="black", linetype="dashed", alpha=0.3, fill="gray55")+
  #  geom_point(aes(mean_sample_elev, female),alpha=0.6, color="darkgreen")+
  # geom_smooth(aes(mean_sample_elev, female), method="lm")+
  #scale_fill_continuous_diverging()+
  theme_cowplot()+
  theme(legend.title = element_blank(),
        legend.position = "none")+
  facet_wrap(.~family, scales="free_y")
```

When we plot log10sex difference (male mass - female mass) by log10midpoint elevation we get a pretty strong negative trend which is in line with the hypothesis that in harsher environments i.e. higher elevations constraints on sexual dimorphism are greater.

## Relationship between sex differences and sexual dimorphism indices
```{r}
p1<-ggplot(df, aes(x=sex_diff, y=SSDI_mass))+
  #geom_point(alpha=0.3)+
  geom_smooth(method="lm", se=T,color="black", linetype="dashed")+
  theme_cowplot()+
  theme(legend.position = "none")

p2<-ggplot(df, aes(x=sex_diff, y=lov.SSDI))+
  #geom_point(alpha=0.3)+
  geom_smooth(method="lm", se=T,color="black", linetype="dashed")+
  theme_cowplot()+
  theme(legend.position = "none")

grid.arrange(p1, p2, nrow=1)
```

This seems to run counter to @lovich1992review, suggesting that (male - female/male) is most comparable?

# Detour to look at only hummingbirds
```{r}
df%>%
  filter(family %in% c("Trochilidae"))%>%
  mutate(mean_mass = (male_mass + female_mass)/2,
         Bee = ifelse(taxon == "Mellisuga helenae", "Bee", NA))%>%
  ggplot(., aes(log10(mean_mass), SSDI_mass, fill=Bee, group=larger_sex))+
  geom_point(shape=21)+
  geom_smooth(method="lm", fill="grey", color="black")+
  geom_abline(intercept = -0.45, slope=1)+
  coord_cartesian(xlim=c(0,1.4), ylim=c(-0.23, 0.6))
```

# SDI Mass
## Elevation
Here I plot SSDI as calculated by (male-female/female)
```{r}
p1 <- df%>%
  ggplot(., aes(x=log10(min_elev), y=SSDI_mass))+
  #geom_point(alpha=0.3)+
  geom_smooth( se=T,color="black", linetype="dashed")+
  theme(legend.position = "none")

p2 <- df%>%
  #filter(family %in% c("Trochilidae", "Furnariidae", "Tyrannidae", "Thraupidae"))%>%
  ggplot(., aes(x=log10(max_elev), y=SSDI_mass))+
  #geom_point(alpha=0.3)+
  geom_smooth( se=T, color="black", linetype="dashed")+
  theme(legend.position = "none")

p3 <- df%>%
  #filter(family %in% c("Trochilidae", "Furnariidae", "Tyrannidae", "Thraupidae"))%>%
  ggplot(., aes(x=log10(mdpt.elev), y=SSDI_mass))+
  #geom_point(alpha=0.3)+
  geom_smooth( se=T, color="black", linetype="dashed")+
  theme(legend.position = "none")

p4 <-ggplot(df, aes(x=log10(amplitude), y=SSDI_mass))+
  #geom_point(alpha=0.3)+
  geom_smooth( se=T, color="black", linetype="dashed")+
  theme(legend.position = "none")

grid.arrange(p1, p2, p3, p4, ncol=2, nrow=2)

```
These look flat when you plot points. There are very slight relationships but, likely nothing to go on.


## Latitude of range
```{r}
p<-ggplot(df, aes(x=Max.Latitude, y=SSDI_mass))+
  #geom_point(alpha=0.3, color="dodgerblue")+
  geom_smooth(method="loess", se=T,color="black", linetype="dashed")+
  theme_cowplot()+
  theme(legend.position = "none")

p1<-ggplot(df, aes(x=Min.Latitude, y=SSDI_mass))+
  #geom_point(alpha=0.3, color="dodgerblue")+
  geom_smooth(method="loess", se=T, color="black", linetype="dashed")+
  theme_cowplot()+
  theme(legend.position = "none")


grid.arrange(p, p1, nrow=1)
```


# SDI Wing
## Elevation

```{r}
p1 <- df%>%
  ggplot(., aes(x=log10(min_elev), y=SSDI_wing))+
  #geom_point(alpha=0.3)+
  geom_smooth( se=T,color="black", linetype="dashed")+
  theme(legend.position = "none")

p2 <- df%>%
  #filter(family %in% c("Trochilidae", "Furnariidae", "Tyrannidae", "Thraupidae"))%>%
  ggplot(., aes(x=log10(max_elev), y=SSDI_wing))+
  #geom_point(alpha=0.3)+
  geom_smooth( se=T, color="black", linetype="dashed")+
  theme(legend.position = "none")

p3 <- df%>%
  #filter(family %in% c("Trochilidae", "Furnariidae", "Tyrannidae", "Thraupidae"))%>%
  ggplot(., aes(x=log10(mdpt.elev), y=SSDI_wing))+
  #geom_point(alpha=0.3)+
  geom_smooth( se=T, color="black", linetype="dashed")+
  theme(legend.position = "none")

p4 <-ggplot(df, aes(x=log10(amplitude), y=SSDI_wing))+
  #geom_point(alpha=0.3)+
  geom_smooth( se=T, color="black", linetype="dashed")+
  theme(legend.position = "none")

grid.arrange(p1, p2, p3, p4, ncol=2, nrow=2)

```

## Latitude
```{r}
p2 <-ggplot(df, aes(x=Max.Latitude, y=SSDI_wing))+
  #geom_point(alpha=0.3, color="dodgerblue")+
  geom_smooth(method="loess", se=T,color="black", linetype="dashed")+
  theme_cowplot()+
  theme(legend.position = "none")

p3 <-ggplot(df, aes(x=Min.Latitude, y=SSDI_wing))+
  #geom_point(alpha=0.3, color="dodgerblue")+
  geom_smooth(method="loess", se=T,color="black", linetype="dashed")+
  theme_cowplot()+
  theme(legend.position = "none")

grid.arrange(p2, p3, ncol=2, nrow=1)

```


# HWI
## Elevation

```{r}

p1 <- df%>%
  ggplot(., aes(x=log10(min_elev), y=`Hand-Wing.Index`))+
  #geom_point(alpha=0.3)+
  geom_smooth( se=T,color="black", linetype="dashed")+
  theme(legend.position = "none")

p2 <- df%>%
  #filter(family %in% c("Trochilidae", "Furnariidae", "Tyrannidae", "Thraupidae"))%>%
  ggplot(., aes(x=log10(max_elev), y=`Hand-Wing.Index`))+
  #geom_point(alpha=0.3)+
  geom_smooth( se=T, color="black", linetype="dashed")+
  theme(legend.position = "none")

p3 <- df%>%
  #filter(family %in% c("Trochilidae", "Furnariidae", "Tyrannidae", "Thraupidae"))%>%
  ggplot(., aes(x=log10(mdpt.elev), y=`Hand-Wing.Index`))+
  #geom_point(alpha=0.3)+
  geom_smooth( se=T, color="black", linetype="dashed")+
  theme(legend.position = "none")

p4 <-ggplot(df, aes(x=log10(amplitude), y=`Hand-Wing.Index`))+
  #geom_point(alpha=0.3)+
  geom_smooth( se=T, color="black", linetype="dashed")+
  theme(legend.position = "none")

grid.arrange(p1, p2, p3, p4, ncol=2, nrow=2)

```

## Latitude
```{r}
p2 <-ggplot(df, aes(x=Max.Latitude, y=`Hand-Wing.Index`))+
  #geom_point(alpha=0.3, color="dodgerblue")+
  geom_smooth(method="loess", se=T,color="black", linetype="dashed")+
  theme_cowplot()+
  theme(legend.position = "none")

p3 <-ggplot(df, aes(x=Min.Latitude, y=`Hand-Wing.Index`))+
  #geom_point(alpha=0.3, color="dodgerblue")+
  geom_smooth(method="loess", se=T,color="black", linetype="dashed")+
  theme_cowplot()+
  theme(legend.position = "none")

grid.arrange(p2, p3, ncol=2, nrow=1)

```

# SDI Blood
## Hemoglobin
```{r}
p<-df%>%
  filter()%>%
ggplot(., aes(max_elev, SSDI_hb, size=Range.Size))+
  geom_point(alpha=0.6,  shape=21)+
  geom_smooth(method="lm", color="black", linetype="dashed")+
  # scale_colour_gradient2(
  #   low = "#d1492d",
  #   mid = "#FFE6BE",
  #   high = "#00798c",
  #   midpoint = 0,
  #   breaks = c(-0.2, 0, 0.2)
  # ) +
  theme_cowplot()+
  coord_cartesian(xlim=c(0,5000), ylim=c(-0.4, 0.5))+
  theme(legend.title = element_blank(),
        legend.position = "none")

p1<-df%>%
  filter()%>%
ggplot(., aes(min_elev, SSDI_hb, size=Range.Size))+
  geom_point(alpha=0.6,  shape=21)+
  geom_smooth(method="lm", color="black", linetype="dashed")+
  # scale_colour_gradient2(
  #   low = "#d1492d",
  #   mid = "#FFE6BE",
  #   high = "#00798c",
  #   midpoint = 0,
  #   breaks = c(-0.2, 0, 0.2)
  # ) +
  theme_cowplot()+
  coord_cartesian(xlim=c(0,5000), ylim=c(-0.4, 0.5))+
  theme(legend.title = element_blank(),
        legend.position = "none")

p2<-df%>%
  filter()%>%
ggplot(., aes(mdpt.elev, SSDI_hb, size=Range.Size))+
  geom_point(alpha=0.6,  shape=21)+
  geom_smooth(method="lm", color="black", linetype="dashed")+
  # scale_colour_gradient2(
  #   low = "#d1492d",
  #   mid = "#FFE6BE",
  #   high = "#00798c",
  #   midpoint = 0,
  #   breaks = c(-0.2, 0, 0.2)
  # ) +
  theme_cowplot()+
  coord_cartesian(xlim=c(0,5000), ylim=c(-0.4, 0.5))+
  theme(legend.title = element_blank(),
        legend.position = "none")

p3<-df%>%
  filter()%>%
ggplot(., aes(amplitude, SSDI_hb, size=Range.Size))+
  geom_point(alpha=0.6,  shape=21)+
  geom_smooth(method="lm", color="black", linetype="dashed")+
  # scale_colour_gradient2(
  #   low = "#d1492d",
  #   mid = "#FFE6BE",
  #   high = "#00798c",
  #   midpoint = 0,
  #   breaks = c(-0.2, 0, 0.2)
  # ) +
  theme_cowplot()+
  coord_cartesian(xlim=c(0,5000), ylim=c(-0.4, 0.5))+
  theme(legend.title = element_blank(),
        legend.position = "none")

grid.arrange(p, p1, p2, p3, ncol=2, nrow=2)
```
Very slight decreasing trend in hemoglobin SDI with maximum elevation.

## Hct
```{r}
p<-df%>%
  filter()%>%
ggplot(., aes(max_elev, SSDI_hct, size=Range.Size))+
  geom_point(alpha=0.6,  shape=21)+
  geom_smooth(method="lm", color="black", linetype="dashed")+
  # scale_colour_gradient2(
  #   low = "#d1492d",
  #   mid = "#FFE6BE",
  #   high = "#00798c",
  #   midpoint = 0,
  #   breaks = c(-0.2, 0, 0.2)
  # ) +
  theme_cowplot()+
  coord_cartesian(xlim=c(0,5000), ylim=c(-0.4, 0.5))+
  theme(legend.title = element_blank(),
        legend.position = "none")

p1<-df%>%
  filter()%>%
ggplot(., aes(min_elev, SSDI_hct, size=Range.Size))+
  geom_point(alpha=0.6,  shape=21)+
  geom_smooth(method="lm", color="black", linetype="dashed")+
  # scale_colour_gradient2(
  #   low = "#d1492d",
  #   mid = "#FFE6BE",
  #   high = "#00798c",
  #   midpoint = 0,
  #   breaks = c(-0.2, 0, 0.2)
  # ) +
  theme_cowplot()+
  coord_cartesian(xlim=c(0,5000), ylim=c(-0.4, 0.5))+
  theme(legend.title = element_blank(),
        legend.position = "none")

p2<-df%>%
  filter()%>%
ggplot(., aes(mdpt.elev, SSDI_hct, size=Range.Size))+
  geom_point(alpha=0.6,  shape=21)+
  geom_smooth(method="lm", color="black", linetype="dashed")+
  # scale_colour_gradient2(
  #   low = "#d1492d",
  #   mid = "#FFE6BE",
  #   high = "#00798c",
  #   midpoint = 0,
  #   breaks = c(-0.2, 0, 0.2)
  # ) +
  theme_cowplot()+
  coord_cartesian(xlim=c(0,5000), ylim=c(-0.4, 0.5))+
  theme(legend.title = element_blank(),
        legend.position = "none")

p3<-df%>%
  filter()%>%
ggplot(., aes(amplitude, SSDI_hct, size=Range.Size))+
  geom_point(alpha=0.6,  shape=21)+
  geom_smooth(method="lm", color="black", linetype="dashed")+
  # scale_colour_gradient2(
  #   low = "#d1492d",
  #   mid = "#FFE6BE",
  #   high = "#00798c",
  #   midpoint = 0,
  #   breaks = c(-0.2, 0, 0.2)
  # ) +
  theme_cowplot()+
  coord_cartesian(xlim=c(0,5000), ylim=c(-0.4, 0.5))+
  theme(legend.title = element_blank(),
        legend.position = "none")

grid.arrange(p, p1, p2, p3, ncol=2, nrow=2)

```

## Hb variability
```{r}
p<-df%>%
  filter()%>%
  mutate(max_elev = ntile(max_elev, 8))%>%
  group_by(max_elev)%>%
  dplyr::summarize(Hb_sd = sd(SSDI_hb, na.rm=T))%>%
ggplot(., aes(max_elev, Hb_sd))+
  geom_point(alpha=0.6,  shape=21)+
  geom_smooth(method="lm", color="black", linetype="dashed")+
  # scale_colour_gradient2(
  #   low = "#d1492d",
  #   mid = "#FFE6BE",
  #   high = "#00798c",
  #   midpoint = 0,
  #   breaks = c(-0.2, 0, 0.2)
  # ) +
  theme_cowplot()+
  theme(legend.title = element_blank(),
        legend.position = "none")

p1<-df%>%
  filter()%>%
  mutate(min_elev = ntile(min_elev, 8))%>%
  group_by(min_elev)%>%
  dplyr::summarize(Hb_sd = sd(SSDI_hb, na.rm=T))%>%
ggplot(., aes(min_elev, Hb_sd))+
  geom_point(alpha=0.6,  shape=21)+
  geom_smooth(method="lm", color="black", linetype="dashed")+
  # scale_colour_gradient2(
  #   low = "#d1492d",
  #   mid = "#FFE6BE",
  #   high = "#00798c",
  #   midpoint = 0,
  #   breaks = c(-0.2, 0, 0.2)
  # ) +
  theme_cowplot()+
  theme(legend.title = element_blank(),
        legend.position = "none")

p2<-df%>%
  filter()%>%
  mutate(mdpt.elev = ntile(mdpt.elev, 8))%>%
  group_by(mdpt.elev)%>%
  dplyr::summarize(Hb_sd = sd(SSDI_hb, na.rm=T))%>%
ggplot(., aes(mdpt.elev, Hb_sd))+
  geom_point(alpha=0.6,  shape=21)+
  geom_smooth(method="lm", color="black", linetype="dashed")+
  # scale_colour_gradient2(
  #   low = "#d1492d",
  #   mid = "#FFE6BE",
  #   high = "#00798c",
  #   midpoint = 0,
  #   breaks = c(-0.2, 0, 0.2)
  # ) +
  theme_cowplot()+
  theme(legend.title = element_blank(),
        legend.position = "none")

p3<-df%>%
  filter()%>%
  mutate(amplitude = ntile(amplitude, 8))%>%
  group_by(amplitude)%>%
  dplyr::summarize(Hb_sd = sd(SSDI_hb, na.rm=T))%>%
ggplot(., aes(amplitude, Hb_sd))+
  geom_point(alpha=0.6,  shape=21)+
  geom_smooth(method="lm", color="black", linetype="dashed")+
  # scale_colour_gradient2(
  #   low = "#d1492d",
  #   mid = "#FFE6BE",
  #   high = "#00798c",
  #   midpoint = 0,
  #   breaks = c(-0.2, 0, 0.2)
  # ) +
  theme_cowplot()+
  theme(legend.title = element_blank(),
        legend.position = "none")

grid.arrange(p, p1, p2, p3, ncol=2, nrow=2)

```

## Hct variability
```{r}
p<-df%>%
  filter()%>%
  mutate(max_elev = ntile(max_elev, 8))%>%
  group_by(max_elev)%>%
  dplyr::summarize(Hct_sd = sd(SSDI_hct, na.rm=T))%>%
ggplot(., aes(max_elev, Hct_sd))+
  geom_point(alpha=0.6,  shape=21)+
  geom_smooth(method="lm", color="black", linetype="dashed")+
  # scale_colour_gradient2(
  #   low = "#d1492d",
  #   mid = "#FFE6BE",
  #   high = "#00798c",
  #   midpoint = 0,
  #   breaks = c(-0.2, 0, 0.2)
  # ) +
  theme_cowplot()+
  theme(legend.title = element_blank(),
        legend.position = "none")

p1<-df%>%
  filter()%>%
  mutate(min_elev = ntile(min_elev, 8))%>%
  group_by(min_elev)%>%
  dplyr::summarize(Hct_sd = sd(SSDI_hct, na.rm=T))%>%
ggplot(., aes(min_elev, Hct_sd))+
  geom_point(alpha=0.6,  shape=21)+
  geom_smooth(method="lm", color="black", linetype="dashed")+
  # scale_colour_gradient2(
  #   low = "#d1492d",
  #   mid = "#FFE6BE",
  #   high = "#00798c",
  #   midpoint = 0,
  #   breaks = c(-0.2, 0, 0.2)
  # ) +
  theme_cowplot()+
  theme(legend.title = element_blank(),
        legend.position = "none")

p2<-df%>%
  filter()%>%
  mutate(mdpt.elev = ntile(mdpt.elev, 8))%>%
  group_by(mdpt.elev)%>%
  dplyr::summarize(Hct_sd = sd(SSDI_hct, na.rm=T))%>%
ggplot(., aes(mdpt.elev, Hct_sd))+
  geom_point(alpha=0.6,  shape=21)+
  geom_smooth(method="lm", color="black", linetype="dashed")+
  # scale_colour_gradient2(
  #   low = "#d1492d",
  #   mid = "#FFE6BE",
  #   high = "#00798c",
  #   midpoint = 0,
  #   breaks = c(-0.2, 0, 0.2)
  # ) +
  theme_cowplot()+
  theme(legend.title = element_blank(),
        legend.position = "none")

p3<-df%>%
  filter()%>%
  mutate(amplitude = ntile(amplitude, 8))%>%
  group_by(amplitude)%>%
  dplyr::summarize(Hct_sd = sd(SSDI_hct, na.rm=T))%>%
ggplot(., aes(amplitude, Hct_sd))+
  geom_point(alpha=0.6,  shape=21)+
  geom_smooth(method="lm", color="black", linetype="dashed")+
  # scale_colour_gradient2(
  #   low = "#d1492d",
  #   mid = "#FFE6BE",
  #   high = "#00798c",
  #   midpoint = 0,
  #   breaks = c(-0.2, 0, 0.2)
  # ) +
  theme_cowplot()+
  theme(legend.title = element_blank(),
        legend.position = "none")

grid.arrange(p, p1, p2, p3, ncol=2, nrow=2)

```

```{r}
p1<-ggplot(df, aes(`Hand-Wing.Index`, SSDI_hb))+
  geom_smooth(method="lm",color="black", linetype="dashed")

p2<-ggplot(df, aes(`Hand-Wing.Index`, SSDI_hct))+
  geom_smooth(method="lm",color="black", linetype="dashed")

grid.arrange(p1, p2, nrow=1)
```
## SSDI Phylo PCA
```{r}

#make matrix

tree <- read.tree("data/birds_mcc.tre")


df2 <- df%>%mutate(taxon = gsub(" ", "_", taxon))%>%filter(taxon %in% pruned.tree$tip.label)%>%dplyr::select(taxon, c(33:34, 49:57))%>%distinct()

tips <- df2%>%
  mutate(taxon = gsub(" ", "_", taxon))%>%
  dplyr::select(taxon)%>%
  pull()

pruned.tree<-drop.tip(tree, tree$tip.label[-na.omit(match(tips, tree$tip.label))])


phy.pca.mat <-scale(as.matrix(df2[ ,c(4:8)]))

rownames(phy.pca.mat) <-df2$taxon
phy.pca.mat<-evobiR::ReorderData(pruned.tree, phy.pca.mat, taxa.names="row names")

bird.sex.pca <-phyl.pca(pruned.tree, phy.pca.mat)
plot(bird.sex.pca)
biplot(bird.sex.pca)

s <-as.data.frame(bird.sex.pca$S)
s$taxon <-names(bird.sex.pca$S)
s <- tibble::rownames_to_column(s, "taxon")
s <-s%>%
   mutate(taxon= gsub("_", " ", taxon))%>%
  left_join(., df, "taxon")%>%
  mutate(min.e.x = scale(min_elev),
         max.e.z = scale(max_elev),
         amp.z = scale(amplitude),
         range.z = scale(Range.Size),
         corrected.PC1 = PC1+4.28)
 

ggplot(s, aes(PC1, SSDI_mass))+
  geom_point(shape=21)

ggplot(s, aes(PC1, SSDI_tarsus))+
  geom_point(shape=21)

ggplot(s, aes(PC1, SSDI_wing))+
  geom_point(shape=21)

ggplot(s, aes(PC1, amplitude))+
  geom_point(shape=21)
  #geom_smooth(method="lm", color="black", linetype="dashed")
  #scale_fill_gradient2(low="blue", mid="yellow", high="red")
```

## Phylo_pca model
```{r}
A <- vcv(pruned.tree)


m1 <- brm(
  formula = bf(PC1 ~ 1 + amp.z*family),
  data = s,
  family = gaussian(), # estimates of  mean and sd in t dist are robust to outliers relative to normal dist
  #data2 = list(A = A), # Phylogeny
  cores = 10,
  chains = 4,
  thin = 10,
  warmup = 10000, # default is iter/2; shouldn't ever be larger than iter
  iter = 20000,
  control = list(adapt_delta = 0.98, max_treedepth = 17),
  sample_prior = TRUE # default priors
)


```

## Model SSDI
### A few large select families
```{r}
tree <- read.tree("data/birds_mcc.tre")
t.brms<-df%>%
  filter(family %in% c("Furnariidae", "Tyrannidae", "Trochilidae", "Thraupidae"))%>%
  mutate(amp.z = scale(amplitude),
         min.z = scale(min_elev),
         max.z =scale(max_elev),
         species = gsub(" ", "_", taxon),
         mdpt.z = scale(mdpt.elev),
         range.size.z =scale(Range.Size))%>%
  filter(species %in% tree$tip.label)

tips <- as.character(t.brms$species)
tree <- keep.tip(tree, tips)
A <- vcv(tree)


m1 <- brm(
  formula = bf(SSDI_mass ~ 1 + s(mdpt.z) +(1|gr(species, cov = A))),
  data = t.brms,
  family = student(), # estimates of  mean and sd in t dist are robust to outliers relative to normal dist
  data2 = list(A = A), # Phylogeny
  cores = 8,
  chains = 4,
  thin = 10,
  warmup = 10000, # default is iter/2; shouldn't ever be larger than iter
  iter = 20000,
  control = list(adapt_delta = 0.98, max_treedepth = 18),
  sample_prior = TRUE # default priors
)
save(m1, file="models/m1_SSDI.RData") # save model
load("models/m2_SSDI.RData") # If loading from pre-saved file and not re-running


```


# Flight Index
## Amplitude
```{r}

#simple tarsus/hand-wing index
df%>%
  mutate(flight_Index = `Hand-Wing.Index`/Tarsus.Length)%>%
  filter(family %in%c("Tyrannidae", "Trochilidae", "Emberizidae", "Furnariidae", "Thraupidae"))%>%
ggplot(., aes(flight_Index, amplitude, fill=Range.Size))+
  geom_point(shape=21, size=3)+
  geom_smooth(method="lm",color="black", linetype="dashed")+
  scale_fill_gradient(low="dodgerblue", high="red3")+
  facet_wrap(.~family, scales = "free")

```
### max_elev
```{r}

#simple tarsus/hand-wing index
df%>%
  mutate(flight_Index = `Hand-Wing.Index`/Tarsus.Length)%>%
  filter(family %in%c("Tyrannidae", "Trochilidae", "Emberizidae", "Furnariidae", "Thraupidae"))%>%
ggplot(., aes(flight_Index, max_elev, fill=amplitude))+
  geom_point(shape=21, size=3)+
  geom_smooth(method="lm",color="black", linetype="dashed")+
  scale_fill_gradient(low="dodgerblue", high="red3")+
  facet_wrap(.~family, scales = "free")

```

# PCA

```{r}
f.pca <- df.all.birds%>%
  mutate(mass= (male+female)/2)%>%
  dplyr::select(species, family, min_elev, max_elev, mass, Kipps.Distance,  Tarsus.Length, `Hand-Wing.Index`, Tail.Length) #pull columns for PCA dataset
f.pca <-na.omit(f.pca) #remove NA's (REALLY SHRINKS DATASET...)
pca.only <-log10(f.pca[,c(5:9)]) #extract morpho characters for PCA
pca.obj <- prcomp(pca.only)#Do PCA on blood characters
f.pca[,c(10:14)]<-pca.obj$x #add PCA's back in to data set for modeling.
setnames(f.pca, old = c('V1', 'V2', 'V3', 'V4', "V5"), new = c('PC1','PC2', 'PC3', 'PC4', "PC5")) #rename PCA value columns

f.pca <- f.pca%>%
  mutate(amplitude= max_elev-min_elev,
         mdpt = (max_elev + min_elev)/2,
         species2 = str_replace(species, " ", "_"),
         amplitude.z = standardize(amplitude),
         max_elev.z = standardize(max_elev))

#scree plot
fviz_eig(pca.obj)

#plot loadings
fviz_pca_var(pca.obj, col.var = "contrib", axes=c(2,3))

f.pca%>%
  filter(PC2 >-2)%>%
  filter(family %in%c("Tyrannidae",  "Emberizidae", "Furnariidae", "Thraupidae")|
         family == "Trochilidae" & PC2>0.6)%>%
ggplot(., aes(PC2,  mdpt, fill=amplitude))+
  geom_point(size=2, shape=21)+
  geom_smooth(method="lm", color="black", linetype="dashed")+
  theme_cowplot()+
  scale_fill_gradient(low="dodgerblue", high="red3")+
  facet_wrap(.~family, scales = "free")

```

## phylogeny for PCA data
```{r}
tyannidae.pca<-f.pca%>%
  filter(family=="Tyrannidae")

tips <- as.character(tyannidae.pca$species2)
tree <- keep.tip(tree, tips)


#what are 7 species not in tree?

f.pca <- f.pca%>%
  filter(!duplicated(species2))

A <- vcv(tree)

#Bayesian models take forever so lets help it by getting pgls for the variables of interest

pic.pca <- pic(f.pca$PC2, tree)
pic.amp <- pic(f.pca$amplitude.z, tree)
pic.max.elev <- pic(f.pca$max_elev.z, tree)

pic.df <- as.data.frame(cbind(pic.pca, pic.amp, pic.max.elev))
```


```{r}
f.m2 <- brm(
  formula = bf(pic.amp ~ 1 + pic.pca * pic.max.elev),
  data = pic.df,
  family = gaussian(), # estimates of  mean and sd in t dist are robust to outliers relative to normal dist
  cores = 4,
  chains = 4,
  thin = 10,
  warmup = 5000, # default is iter/2; shouldn't ever be larger than iter
  iter = 10000,
  control = list(adapt_delta = 0.99, max_treedepth = 15),
  sample_prior = TRUE # default priors
)
save(hb.m8, file="hb.m8_ReducedModel_PhyloREOnly.RData") # save model
load("hb.m8_ReducedModel_PhyloREOnly.RData") # If loading from pre-saved file and not re-running
conditional_effects(f.m2)
```

All of this is before correcting for phyologeny which we should be next...

## Modeling
```{r}
#read in big tree 
#tree <- read.tree("data/tree_3432spp.tre")
A <- vcv(tree)

df.all.birds <-df.all.birds%>%
  mutate(max_elev.z = standardize(max_elev),
         species2 = as.character(str_replace(species, " ", "_")))

tyannidae.pca <-tyannidae.pca%>%
  mutate(mdp.z =scale(mdpt),
         amp)

get_prior(PC2 ~ 1 + mdp.z + (1|gr(species2, cov=A)), data=tyannidae.pca, data2=list(A=A))
#MODEL 2: PC2 ~ mdpt PHYLO RANDOM EFFECT
m1 <- brm(
  formula = bf(PC2 ~ 1 + mdp.z*amplitude.z + (1|gr(species2, cov=A))),
  data = tyannidae.pca,
  family = student(), # estimates of  mean and sd in t dist are robust to outliers relative to normal dist
  data2 = list(A = A), # Phylogeny
  cores = 6,
  chains = 4,
  thin = 10,
  warmup = 10000, # default is iter/2; shouldn't ever be larger than iter
  iter = 20000,
  control = list(adapt_delta = 0.99, max_treedepth = 17),
  sample_prior = TRUE # default priors
)
save(ssdi.m2, file="models/m2_SSDI.RData") # save model
load("models/m2_SSDI.RData") # If loading from pre-saved file and not re-running



```

### Checking

```{r}
hist(df$SSDI)
#not bad, some long tails...
plot(tree, type="fan", cex=0.1)

```


```{r}
df.cont <- df|>
  mutate(species = as.character(str_replace(Species, " ", "_")))|>
  filter(species %in% tree$tip.label)|>
  ungroup()|>
  dplyr::select(SSDI_mass)|>
  na.omit()%>%
  pull()

tips <- df|>
  mutate(species = as.character(str_replace(Species, " ", "_")))|>
  filter(species %in% tree$tip.label)|>
  ungroup()|>
  dplyr::select(species, SSDI_mass)|>
    na.omit()%>%
  pull(species)

tree <- keep.tip(tree, tips)

names(df.cont) <- tree$tip.label

fitEB<-anc.ML(tree,df.cont,model="EB")

obj<-contMap(tree, df.cont)
obj<-setMap(obj,colors=c("#d1492d","#FFE6BE","#00798c"))
plot(obj, fsize=c(0.1,1), outline=F, lwd=1.3, type="fan")

```