---
title: "Macro SSDI"
author: "Chauncey Gadek" 
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    number_sections: true
    toc_depth: 5
    code_folding: show
    bibliography: true
    #df_print: paged
    #df_print: kable
    #toc_float: true
      #collapsed: false
      #smooth_scroll: TRUE
    theme: cosmo #spacelab #yeti #united #cosmo
    highlight: tango
  pdf_document:
    df_print: kable
fontsize: 12pt
geometry: margin=0.25in
always_allow_html: yes
---

```{=html}
<style>
/* HTML FORMATTING */
h1, .h1, h2, .h2, h3, .h3, h4, .h4, h5, .h5 {
  margin-top: 25px; /* space before each header */
  font-weight: bold; /* bold headers */
}
</style>
```
```{R, echo=FALSE}
# I set some GLOBAL R chunk options here.
#   (to hide this message add "echo=FALSE" to the code chunk options)
rm(list =ls (all = TRUE)) #This removes objects from global environ
knitr::opts_chunk$set(echo=F, comment = NA, message = FALSE, warning = FALSE, width = 100)
knitr::opts_chunk$set(fig.align = "center", fig.height = 4, fig.width = 6)
#setwd("C:/Users/Trinity's/Dropbox/Crane_trachea")
#setwd("~/Dropbox/Crane_trachea") #need to change to local directory
```

# Load Packages

```{R, echo=F}
library(reshape)
library(reshape2)
library(plyr)
library(dplyr)
library(tidyr)
library(car)
library(rcompanion)
library(GGally)
library(Hmisc)
library(gridExtra)
library(stats)
library(gplots)
library(ggplot2)
library(ggExtra)
library(cowplot)
library(colorspace)
library(stats4) # Forces knitr to work when it's being wonky
library(PMCMR) #Allows Kruskal-Wallis post-hocs
library(effects)
library(gridExtra)
library(lattice)
library(survival)
library(fmsb)
library(faraway)
library(ape)
#library(wBoot)
library(ggridges)
library(boot)
library(faux)
library(effsize)
library(plotrix)
library(colorspace)
library(ggpubr)
library(patchwork)
library(ggdist)

# Mapping 
library(raster)
library(sp)
library(rgdal)
library(RStoolbox)
library(prettymapr)
library(viridis)
library(rasterVis)


# Modeling packages 
library(nlme)
library(lme4)
library(AICcmodavg)
library(MuMIn)
library(glmulti)
library(reghelper)
library(lsmeans)
library(rsq) # get r-squared values from GLM
library(r2glmm) # for R^2 values from lmer() and glmer()
library(multcompView) # related to multiple comparisons?
library(jtools) # interaction plots 
library(interactions) # interaction plots 
library(broom)
library(stargazer) # model output tables
library(ggeffects) # for estimating model predictions from mixed effects models
library(MCMCglmm)
library(bayesplot)
library(rstan)
library(Rcpp) # required for brms
library(brms)
library(magrittr)
library(tidybayes)
library(modelr)
library(hexbin)
library(ggExtra)
library(rgl)
library(readr)
library(tidyverse)

# To run each time you load rstan
options(mc.cores = parallel::detectCores()) # for core setup 
rstan_options(auto_write = TRUE) # auto save  bare verion of compiled Stan program to HD

# Phylo packages 
library(phytools)
library(ape)

#Load in functions
source("~/Dropbox/Gadek_custom_colors.R")



#setup folder paths for less coding
figures <- paste(getwd(), "/figures/", sep="")
tables <- paste(getwd(), "/Tables/", sep="")
models <- paste(getwd(), "/models/", sep="")
results <- paste(getwd(), "/models/results/", sep="")
```

Sexual size dimorphism is a long-studied phenomenon in animals. In this exploration, I examine a few historical methods of effectively comparing and scaling SSD between species @lovich1992review, and I extend studies of sexual dimorphism across elevation gradients, mainly in insects,(@laiolo2013local, @bowden2013fecundity), to Andean birds using phylogenetic mixed models.

The debate about the proper metric for comparing sexual size dimorphism among taxa has yielded one heuristic: simpler is better. @lovich1992review reviewed various ratio metrics of sexual size dimorphism looking at scalability and intuitive utility and recommend: 

$$
SDI = \frac{size of larger sex}{size of smaller sex}
$$


# Data wrangling

I'm mainly using the dataset used in @linck2021blood downloaded from Ethan Linck's [Github repo](https://github.com/elinck/andean_range_limits). I am supplementing this with elevational range data collated from @del1992handbook and avian trait data downloaded from the [Open Traits Network](https://opentraits.org/datasets/avonet), (@tobias2022avonet).

```{r}

EG_Witt <- read_csv("~/Dropbox/Research/Marsh Birds/Marsh_birds/data/EG_Witt.csv")

Linck_et_al_data <- read_csv("~/Dropbox/Research/Marsh Birds/Marsh_birds/data/Linck_et-al_data.csv")

elevs <- read_csv("~/Dropbox/Research/Macro_SSDI/data/elevs.csv")

avonet <- read_csv("~/Dropbox/Research/Avian_Evo_Trans_Rev/analysis/data/Birdtree_AVONET.csv")

#read in tree file
#tree <- read.tree("~/Dropbox/Research/Marsh Birds/Marsh_birds/data/BirdzillaHackett10.tre")

avonet <- avonet%>%
  mutate(species = factor(Species3),
         family = factor(Family3))%>%
  dplyr::select(species, family, Beak.Length_Culmen, Beak.Length_Nares, Beak.Width, Beak.Depth, Tarsus.Length, Kipps.Distance, `Hand-Wing.Index`, Tail.Length, Habitat, Habitat.Density, Migration, Trophic.Level, Trophic.Niche, Min.Latitude, Max.Latitude, Centroid.Latitude, Centroid.Longitude, Range.Size)

elevs <- elevs%>%
  mutate(species = str_replace(species, "_", " "),
         species = factor(species))|>
  distinct()

df <-Linck_et_al_data%>%
  filter(!is.na(mass),
         sex %in% c("male", "female"))%>%
  mutate(elev_bin = cut(elevation, breaks=7),
         sex = factor(sex),
         species = factor(species)
         )%>%
     group_by(species) %>% 
  filter(all(levels(sex) %in% sex))%>%
  group_by(species, sex) %>% 
  summarise(mean_mass = mean(mass),
            )%>%
  pivot_wider(id_cols = c(species), names_from = c(sex), values_from = c(mean_mass))%>%
  mutate(SSDI = (male-female)/female)

L.elevs<- Linck_et_al_data%>%
  group_by(species)%>%
  summarise(max_sample_elev = max(elevation, na.rm=T),
            mean_sample_elev = mean(elevation, na.rm=T),
            min_sample_elev = min(elevation, na.rm=T))
 

df<-df%>%
  left_join(., L.elevs, by="species")

df<-df%>%
  left_join(., elevs, by = "species")

df<- df%>%
  left_join(., avonet, by = "species")%>%
   mutate(amplitude = max_elev - min_elev)

#generate Lovich's simple SSDI Metric

df <- df%>%
  mutate(lov.SSDI = if_else(male > female, (male/female), (female/male)),
         sex_diff = male-female,
         species = factor(species))%>%
  na.omit()

#get vector of species for tree
species <- df|>
  dplyr::select(species)|>
  mutate(species = as.character(str_replace(species, " ", "_")))|>
  pull()

tree.t <- as.phylo(tree[[sample(1:1000, size=1)]])

tree <- keep.tip(tree.t, tip = species)

write.tree(tree, file="~/Dropbox/Research/Marsh Birds/Marsh_birds/data/tree.tre")

#get variance covariance matrix for models
A <-ape::vcv(tree)

```

# Explore data
## Relationship between sex differences and sexual dimorphism indices
```{r}
p1<-ggplot(df, aes(x=sex_diff, y=SSDI))+
  #geom_point(alpha=0.3)+
  geom_smooth(method="lm", se=T)+
  theme_cowplot()+
  theme(legend.position = "none")

p2<-ggplot(df, aes(x=sex_diff, y=lov.SSDI))+
  #geom_point(alpha=0.3)+
  geom_smooth(method="lm", se=T)+
  theme_cowplot()+
  theme(legend.position = "none")

grid.arrange(p1, p2, nrow=1)
```

This seems to run counter to @lovich1992review, suggesting that (male -female/male) is most comparable?

## SSDI x elevation

Here I plot SSDI as calculated by (male-female/female)
```{r}

p1<-ggplot(df, aes(x=min_elev, y=SSDI))+
  #geom_point(alpha=0.3)+
  geom_smooth(method="lm", se=T)+
  theme_cowplot()+
  theme(legend.position = "none")

p2 <-ggplot(df, aes(x=max_elev, y=SSDI))+
  #geom_point(alpha=0.3)+
  geom_smooth(method="lm", se=T)+
  theme_cowplot()+
  theme(legend.position = "none")

p3 <-ggplot(df, aes(x=mean_sample_elev, y=SSDI))+
  #geom_point(alpha=0.3)+
  geom_smooth(method="lm", se=T)+
  theme_cowplot()+
  theme(legend.position = "none")

p4 <-ggplot(df, aes(x=amplitude, y=SSDI))+
  #geom_point(alpha=0.3)+
  geom_smooth(method="lm", se=T)+
  theme_cowplot()+
  theme(legend.position = "none")

grid.arrange(p1, p2, p3, p4, ncol=2, nrow=2)

```

Here I plot SSDI as calculated in @llovich1992review (mass larger sex/mass smaller sex)
```{r}

p1<-ggplot(df, aes(x=min_elev, y=lov.SSDI))+
  #geom_point(alpha=0.3)+
  geom_smooth(method="lm", se=T)+
  theme_cowplot()+
  theme(legend.position = "none")

p2 <-ggplot(df, aes(x=max_elev, y=lov.SSDI))+
  #geom_point(alpha=0.3)+
  geom_smooth(method="lm", se=T)+
  theme_cowplot()+
  theme(legend.position = "none")

p3 <-ggplot(df, aes(x=mean_sample_elev, y=lov.SSDI))+
  #geom_point(alpha=0.3)+
  geom_smooth(method="lm", se=T)+
  theme_cowplot()+
  theme(legend.position = "none")

p4 <-ggplot(df, aes(x=amplitude, y=lov.SSDI))+
  #geom_point(alpha=0.3)+
  geom_smooth(method="lm", se=T)+
  theme_cowplot()+
  theme(legend.position = "none")

grid.arrange(p1, p2, p3, p4, ncol=2, nrow=2)

```

Here we see the sexual dimporphism index clearly goes down... More in line with my predictions that sexual dimorphism will decrease as abiotic selection increases with elevation...

## Latitude of range
```{r}
p1<-ggplot(df, aes(x=Max.Latitude, y=lov.SSDI))+
  #geom_point(alpha=0.3)+
  geom_smooth(method="lm", se=T)+
  theme_cowplot()+
  theme(legend.position = "none")

p2 <-ggplot(df, aes(x=max_elev, y=Max.Latitude))+
  #geom_point(alpha=0.3)+
  geom_smooth(method="lm", se=T)+
  theme_cowplot()+
  theme(legend.position = "none")

p3 <-ggplot(df, aes(x=mean_sample_elev, y=Max.Latitude))+
  geom_point(alpha=0.3)+
  geom_smooth(method="lm", se=T)+
  theme_cowplot()+
  theme(legend.position = "none")


```



All of this is before correcting for phylogeny which we should be next...

## Modeling
### Checking


```{r}
hist(df$SSDI)
#not bad, some long tails...

```
